% -------------------------------------------------------------------------------------------------
% --------------------------------------- Acknowledgements ----------------------------------------
% -------------------------------------------------------------------------------------------------
\begin{titlepage}
%----------------------------------------------
% Acknowledgements Header
\begin{center}
{\huge \bfseries Acknowledgements}\\[1.5cm]
\end{center}
%----------------------------------------------
% Text for Acknowledgements
I would like to offer my sincere thanks to Dr M.S. Madhusudhan, Associate Professor at the Indian Institute of Science Education and Research, Pune for providing me with the opportunity to work on this project. His invaluable guidance at each and every step of the project helped me develop a scientific outlook and his high energy levels kept me enthused throughout the term of the project and thereafter. I thoroughly enjoyed the countless debugging and brainstorming sessions that we had together. Thanks for being a great mentor.
\par
I would also like to express my gratitude to Dr. Richa Rikhy, Assistant Professor at IISER Pune for the helpful comments she provided during the course of the project. Also, thanks for agreeing to be on my Thesis Advisory Committee.
\par
I thank all the young researchers in the Madhusudhan Lab, Neelesh Soni, Sagar Gore, Kuan Pern Tan, Yogendra Ramtirtha, Parichit Sharma, Shipra Gupta, Neeladri Sen for the valuable discussions and support and also for the numerous cricket and dart games that we enjoyed together.
\par
It would be wrong if I do not thank my friends at IISER Pune, who have provided me with fun and joy and have been very supportive (for all the good and the not so good deeds). Thanks for the countless memories we made together.
\par
Finally, I wholeheartedly thank my parents and the rest of the family, who have been my true support. I thank them for believing in me and letting my follow my heart and study the natural world.
\par
I would also like to thank Kalyanee Shirlekar and Sharvaree Vadgama for the beers they promised, given I put their names here.
%----------------------------------------------
\vfill % Fills remaining of the vertical space with blank space
%----------------------------------------------
\end{titlepage}
% -------------------------------------------------------------------------------------------------
% --------------------------------------- Acknowledgements ----------------------------------------
% -------------------------------------------------------------------------------------------------